### 2018-01-12 submission/manuscript.Rmd
add figure legends and update discussion and results

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Using PERMANOVA to test whether there were any differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Interestingly, Phusion was one of only two DNA polymerase that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5 cycle increments within each sample. We found that there was no detectable difference in Bray-Curits index when comparing to the previous 5 cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curits indicies disappear for both differences in cycle number and within 5 cycle increments within an individual.

***Sequence Error is Dependent on both Taq and Cycle Number Used:***
Differences by HiFi Taq in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest values were for the Kappa HiFI DNA polymerase [Figure 3A]. Sub-sampling depth seems to have little effect on this rate with both 5000 and 10000 sub-sampled sequences showing similar results [Figure 3B-C]. There were small differences between the various HiFi Taq at lower cycle number but larger differences at higher cycle number with Platinum having large differences between all other HiFi DNA polymerase [Figure 3B-C and Table S6].

@@

***Chimeric Sequences Corelate with OTUs and are HiFi Taq Dependent:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Using a Dunn's post-hoc test the vast majority of these differences were driven by Platinum being different then all other HiFi DNA polymerase across cycle number and sub-sampling depth [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerase [Figure 4].

For all Taqs a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Platinum and Phusion HiFi DNA Polymerase [Figure 5]. In general, the R^2^ value between the number of OTUs and chimeric sequences became stronger as sub-sampling depth increased [Figure 5]. Taken together this data suggests that a strong correlation exists between the number of OTUs and chimera sequence prevalence. Interestingly, Accuprime^TM^ not only had one of the lowest prevalence of chimeric sequences but also, consistently, had the lowest correlations between the number of OTUs and chimeric sequences across sub-sampling depth [Figure 5].

@@

Although we did not observe strong differences, based on cycle number, using the Bray-Curtis index the data suggests that there may be differences between 15 cycles and higher cycle numbers, such as 30, that are commonly used. What is most interesting is that there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). This finding, in conjunction with the PERMANOVA results, would suggest that differences based on Bray-Curtis incrementaly build up so that within 5 cycle increments the differences are not that large but when comparing larger cycle differences, as with the PERMANOVA, these differences start to become detectable. Although our data suggests that it is possible to overcome these types of cycle dependent differences with increased sub-sampling depth, for some DNA polymerase increased sub-sampling depth increases chimera prevalence and per base error rate. Conversely, it is possible that many of the increased number of OTUs, generated as cycle number increases, are not highly abundant and the Bray-Curtis index is able to successfully downweight these respective OTUs [@bc_index_Minchin1987].

@@

@@

**Table 1: **

\footnotesize

```{r table1, echo=FALSE, message=FALSE, warning=FALSE}

\normalsize
\newpage

**Table 2: **

\footnotesize

```{r table2, echo=FALSE, message=FALSE, warning=FALSE}

\normalsize
\newpage

**Figure 1: .**

**Figure 2: .**

**Figure 3: .**

**Figure 4: .**

**Figure 5: .**

**Figure 6: .**

**Figure 7: .**

@@

**Figure S1: .**

**Figure S2: .**

**Figure S3: .**

**Figure S4: .**

**Figure S5: .**

**Figure S6: .**

**Figure S7: .**

>>>>

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Overall, there was very little difference between each respective 5-cycle increment (e.g. 15x vs 20x) for both fecal and mock samples and this was consistent across the different sub-samplings used [Figure 3]. Two exceptions to the low differences between 5-cycle increments can be found. The first that there seems to be large differences for fecal samples between 20x vs. 25x that is robust against sub-sampling depth [Figure 3A-B]. Second, for the mocks, where data is available, there appears to be a similar large difference detected between 20x and 25x [Figure 3D]. Regardless, by the time PCR cycles reach 25x there does not seem to be large differences in the community [Figure 3].

Using PERMANOVA to test whether there were any differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Interestingly, Phusion was one of only two DNA polymerase that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5 cycle increments within each sample. We found that there was no detectable difference in Bray-Curits index when comparing to the previous 5 cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curits indicies disappear for both differences in cycle number and within 5 cycle increments within an individual.

***Sequence Error is Dependent on both Taq and Cycle Number Used:***
Differences by HiFi Taq in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest values were for the Kappa HiFI DNA polymerase [Figure 4A]. Sub-sampling depth seems to have little effect on this rate with both 5000 and 10000 sub-sampled sequences showing similar results [Figure 4B-C]. There were small differences between the various HiFi Taq at lower cycle number but larger differences at higher cycle number with Platinum having large differences between all other HiFi DNA polymerase [Figure 4B-C and Table S6].

@@

***Chimeric Sequences Corelate with OTUs and are HiFi Taq Dependent:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Using a Dunn's post-hoc test the vast majority of these differences were driven by Platinum being different then all other HiFi DNA polymerase across cycle number and sub-sampling depth [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerase regardless of whether pre.cluster or VSEARCH had been used [Figure 5].

For all Taqs a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. In general, the R^2^ value between the number of OTUs and chimeric sequences did not change from the use of pre.cluster and VSEARCH [Figure 6]. Taken together this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences. Kappa had the highest per base error rate and the lowest correlations between the number of OTUs and chimeric sequences across sub-sampling depth [Figure 4 & 6].

@@

Although we did not observe strong differences, based on cycle number, using the Bray-Curtis index the data suggests that there may be differences between 15 cycles and higher cycle numbers, such as 30x, that are commonly used. What is most interesting is that there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). Conversely, this may be due to low power and on observation there does seem to be a trend that 20x and 25x communities are very different [Figure 3]. This finding, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once at least 25 cylces are reached. Increasing the sub-sampling depth, for some DNA polymerase may reduce some of these observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence differences between the different HiFi DNA polymerases [Figure 5]. The chimera prevalence was very strongly correlated with the number of OTUs and this value is relied upon heavily for many different downstream community metric calculations. However, Bray-Curtis analysis with PERMANOVA showed few differences based on DNA polymerase. Since it is possible that many of the increased number of OTUs, generated as cycle number increases, are not highly abundant that the Bray-Curtis index is able to successfully downweight these respective OTUs [@bc_index_Minchin1987]. So, choice of downstream diversity metric could be an important consideration to mitigate these observed changes due to high chimera prevalence in HiFi DNA polymerase such as Platinum.

Our observations suggest that there are clear HiFi dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), it is not clear what the difference between the two different mixes really is. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate since all samples were also sequenced at the same time as well as amplified using the same machine.

@@

Our findings show that measures that rely on number of OTUs will be specific for a particular study and may not be easily generalized to other studies that may be studying a similar area. Care should be taken when choosing a HiFi DNA polymerase for 16S rRNA gene surveys since intrinsic differences can change the number of OTUs observed as well as potentially influence diversity based metrics that do not down weight rare observations.

@@

**Figure 1: Normalized Fecal Number of OTUs.** A) Sub-sampled to 1000 reads. B) Sub-sampled to 5000 reads. C) Sub-sampled to 10000 reads. D) Sub-sampled to 15000 reads. The dotted line represents no change from the mean number of OTUs within that specific individual.

**Figure 2:  Mock Sample Variability in Number of OTUs based on HiFi DNA Polymerase.** A) Sub-sampled to 1000 reads. B) Sub-sampled to 5000 reads. C) Sub-sampled to 10000 reads. The dotted line represents the number of OTUs genereated when the mock reference sequences are run through the pipeline.

**Figure 3: Five Cycle Interval Community Differences.** A) Fecal samples sub-sampled to 1000 reads. B) Fecal samples sub-sampled to 5000 reads. C) Fecal samples sub-sampled to 10000 reads. D) Mock samples sub-sampled to 1000 reads. E) Mock samples sub-sampled to 5000 reads. F) Mock samples sub-sampled to 10000 reads. The solid black lines represent the median Bray-Curtis index difference within sample for each 5 cycle interval.

**Figure 4: HiFi DNA Polymerase Per Base Error Rate in Mock Samples.** A) Error rate before the merger of sequences with pre.cluster and the removal of chimeras with VSEARCH. B) Error rate before the removal of chimeras with VSEARCH. C) Full pipeline. The error bars represent the 75% interquartile range of the median.

**Figure 5: HiFi DNA Polymerase Chimera Prevalence in Mock Samples.** A) Chimera sequence percentage before the merger of sequences with pre.cluster and the removal of chimeras with VSEARCH. B) Chimera sequence percentage before the removal of chimeras with VSEARCH. C) Full pipeline. The error bars represent the 75% interquartile range of the median.

**Figure 6: The Correlation between Number of OTUs and Chimeras.** A) Correlation before the merger of sequences with pre.cluster and the removal of chimeras with VSEARCH. B) Correlation before the removal of chimeras with VSEARCH. C) Correlation with full pipeline.

@@

**Figure S1: HiFi DNA Polymerase Sequence Error Prevalence in Mock Samples.** A) Sequence error prevalence before the merger of sequences with pre.cluster and the removal of chimeras with VSEARCH. B) Sequence error prevalence before the removal of chimeras with VSEARCH. C) Full pipeline. The error bars represent the 75% interquartile range of the median.

**Figure S2: HiFi DNA Polymerase Nucleotide Subsitutions in Mock Samples.**

### 2018-01-12 submission/manuscript.tex
add figure legends and update discussion and results

\textbf{\emph{Statistical Analysis:}} All analysis was done with the R
(v 3.4.2) software package (R Core Team, 2017). Data tranformation and
graphing was completed using the tidyverse package (v 1.1.1) and colors
selected using the viridis package (v 0.4.0) (Garnier, 2017; Wickham,
2017). The total number of OTUs were analyzed using an ANOVA with a
tukey post-hoc test. For the fecal samples the data was normalized to
each individual by cycle number to account for the biological variation
between different people. Bray-Curtis matrices were generated using
mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000. The
distance matrix data was analyzed using PERMANOVA with the vegan package
(v 2.4.4) (Oksanen et al., 2017) and kruskal-wallis tests within R. For
both error and chimera analysis, samples were tested using
Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction
for multiple comparison utilized the Benjamini-Hochberg method
(Benjamini \& Hochberg, 1995).

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Using PERMANOVA to test whether there were any differences
within HiFi DNA polymerase groups based on cycle number, only Phusion
had cycle dependent differences at 1000 and 5000 sub-sampling depth
(P-value = 0.03 and 0.01, respectively). Interestingly, Phusion was one
of only two DNA polymerase that managed to have samples for the 1000
sub-sampling depth at 15 cycles. Next, we assessed whether there were
any major differences between 5 cycle increments within each sample. We
found that there was no detectable difference in Bray-Curits index when
comparing to the previous 5 cycle increment (P-value \textgreater{}
0.05). However, Phusion at 1000 sub-sampling depth had a P-value = 0.02
before multiple comparison correction. It should be noted that at higher
sub-sampling depths these differences in Bray-Curits indicies disappear
for both differences in cycle number and within 5 cycle increments
within an individual.

\textbf{\emph{Sequence Error is Dependent on both Taq and Cycle Number
Used:}} Differences by HiFi Taq in the median average per base error
varied without a clear pattern across sub-sampling depth {[}Table S5{]}.
Generally, the highest values were for the Kappa HiFI DNA polymerase
{[}Figure 3A{]}. Sub-sampling depth seems to have little effect on this
rate with both 5000 and 10000 sub-sampled sequences showing similar
results {[}Figure 3B-C{]}. There were small differences between the
various HiFi Taq at lower cycle number but larger differences at higher
cycle number with Platinum having large differences between all other
HiFi DNA polymerase {[}Figure 3B-C and Table S6{]}.

@@

\textbf{\emph{Chimeric Sequences Corelate with OTUs and are HiFi Taq
Dependent:}} After chimera removal using VSEARCH and removal of
sequences that did not classify as bacteria we assessed the percentage
of sequences that were still chimeric within our mock community. At all
levels of sub-sampling and cycle number there were significant
differences between the HiFi DNA polymerase used (P-value \textless{}
0.05) {[}Table S9{]}. Using a Dunn's post-hoc test the vast majority of
these differences were driven by Platinum being different then all other
HiFi DNA polymerase across cycle number and sub-sampling depth {[}Table
S10{]}. Generally, across sub-sampling depth and cycle number
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
the HiFi DNA Polymerase {[}Figure 4{]}.

For all Taqs a positive correlation was observed between chimeric
sequences and number of OTUs, with this correlation being strongest for
Platinum and Phusion HiFi DNA Polymerase {[}Figure 5{]}. In general, the
R\textsuperscript{2} value between the number of OTUs and chimeric
sequences became stronger as sub-sampling depth increased {[}Figure
5{]}. Taken together this data suggests that a strong correlation exists
between the number of OTUs and chimera sequence prevalence.
Interestingly, Accuprime\textsuperscript{TM} not only had one of the
lowest prevalence of chimeric sequences but also, consistently, had the
lowest correlations between the number of OTUs and chimeric sequences
across sub-sampling depth {[}Figure 5{]}.

@@

Although we did not observe strong differences, based on cycle number,
using the Bray-Curtis index the data suggests that there may be
differences between 15 cycles and higher cycle numbers, such as 30, that
are commonly used. What is most interesting is that there was no
difference within individuals between corresponding 5 cycle increments
(e.g.~15 to 20, 20 to 25, etc.). This finding, in conjunction with the
PERMANOVA results, would suggest that differences based on Bray-Curtis
incrementaly build up so that within 5 cycle increments the differences
are not that large but when comparing larger cycle differences, as with
the PERMANOVA, these differences start to become detectable. Although
our data suggests that it is possible to overcome these types of cycle
dependent differences with increased sub-sampling depth, for some DNA
polymerase increased sub-sampling depth increases chimera prevalence and
per base error rate. Conversely, it is possible that many of the
increased number of OTUs, generated as cycle number increases, are not
highly abundant and the Bray-Curtis index is able to successfully
downweight these respective OTUs (Minchin, 1987).

@@

@@

\textbf{Table 1: }

\footnotesize

\normalsize
\newpage

\textbf{Table 2: }

\footnotesize

\normalsize
\newpage

\textbf{Figure 1: .}

\textbf{Figure 2: .}

\textbf{Figure 3: .}

\textbf{Figure 4: .}

\textbf{Figure 5: .}

\textbf{Figure 6: .}

\textbf{Figure 7: .}

@@

\textbf{Figure S1: .}

\textbf{Figure S2: .}

\textbf{Figure S3: .}

\textbf{Figure S4: .}

\textbf{Figure S5: .}

\textbf{Figure S6: .}

\textbf{Figure S7: .}

>>>>

\textbf{\emph{Statistical Analysis:}} All analysis was done with the R
(v 3.4.3) software package (R Core Team, 2017). Data tranformation and
graphing was completed using the tidyverse package (v 1.1.1) and colors
selected using the viridis package (v 0.4.0) (Garnier, 2017; Wickham,
2017). The total number of OTUs were analyzed using an ANOVA with a
tukey post-hoc test. For the fecal samples the data was normalized to
each individual by cycle number to account for the biological variation
between different people. Bray-Curtis matrices were generated using
mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000. The
distance matrix data was analyzed using PERMANOVA with the vegan package
(v 2.4.4) (Oksanen et al., 2017) and kruskal-wallis tests within R. For
both error and chimera analysis, samples were tested using
Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction
for multiple comparison utilized the Benjamini-Hochberg method
(Benjamini \& Hochberg, 1995).

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Overall, there was very little difference between each
respective 5-cycle increment (e.g.~15x vs 20x) for both fecal and mock
samples and this was consistent across the different sub-samplings used
{[}Figure 3{]}. Two exceptions to the low differences between 5-cycle
increments can be found. The first that there seems to be large
differences for fecal samples between 20x vs.~25x that is robust against
sub-sampling depth {[}Figure 3A-B{]}. Second, for the mocks, where data
is available, there appears to be a similar large difference detected
between 20x and 25x {[}Figure 3D{]}. Regardless, by the time PCR cycles
reach 25x there does not seem to be large differences in the community
{[}Figure 3{]}.

Using PERMANOVA to test whether there were any differences within HiFi
DNA polymerase groups based on cycle number, only Phusion had cycle
dependent differences at 1000 and 5000 sub-sampling depth (P-value =
0.03 and 0.01, respectively). Interestingly, Phusion was one of only two
DNA polymerase that managed to have samples for the 1000 sub-sampling
depth at 15 cycles. Next, we assessed whether there were any major
differences between 5 cycle increments within each sample. We found that
there was no detectable difference in Bray-Curits index when comparing
to the previous 5 cycle increment (P-value \textgreater{} 0.05).
However, Phusion at 1000 sub-sampling depth had a P-value = 0.02 before
multiple comparison correction. It should be noted that at higher
sub-sampling depths these differences in Bray-Curits indicies disappear
for both differences in cycle number and within 5 cycle increments
within an individual.

\textbf{\emph{Sequence Error is Dependent on both Taq and Cycle Number
Used:}} Differences by HiFi Taq in the median average per base error
varied without a clear pattern across sub-sampling depth {[}Table S5{]}.
Generally, the highest values were for the Kappa HiFI DNA polymerase
{[}Figure 4A{]}. Sub-sampling depth seems to have little effect on this
rate with both 5000 and 10000 sub-sampled sequences showing similar
results {[}Figure 4B-C{]}. There were small differences between the
various HiFi Taq at lower cycle number but larger differences at higher
cycle number with Platinum having large differences between all other
HiFi DNA polymerase {[}Figure 4B-C and Table S6{]}.

@@

\textbf{\emph{Chimeric Sequences Corelate with OTUs and are HiFi Taq
Dependent:}} After chimera removal using VSEARCH and removal of
sequences that did not classify as bacteria we assessed the percentage
of sequences that were still chimeric within our mock community. At all
levels of sub-sampling and cycle number there were significant
differences between the HiFi DNA polymerase used (P-value \textless{}
0.05) {[}Table S9{]}. Using a Dunn's post-hoc test the vast majority of
these differences were driven by Platinum being different then all other
HiFi DNA polymerase across cycle number and sub-sampling depth {[}Table
S10{]}. Generally, across sub-sampling depth and cycle number
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
the HiFi DNA Polymerase regardless of whether pre.cluster or VSEARCH had
been used {[}Figure 5{]}.

For all Taqs a positive correlation was observed between chimeric
sequences and number of OTUs, with this correlation being strongest for
Accuprime, Platinum and Phusion HiFi DNA Polymerase {[}Figure 6{]}. In
general, the R\textsuperscript{2} value between the number of OTUs and
chimeric sequences did not change from the use of pre.cluster and
VSEARCH {[}Figure 6{]}. Taken together this data suggests that a strong
correlation exists between the number of OTUs and the prevalence of
chimeric sequences. Kappa had the highest per base error rate and the
lowest correlations between the number of OTUs and chimeric sequences
across sub-sampling depth {[}Figure 4 \& 6{]}.

@@

Although we did not observe strong differences, based on cycle number,
using the Bray-Curtis index the data suggests that there may be
differences between 15 cycles and higher cycle numbers, such as 30x,
that are commonly used. What is most interesting is that there was no
difference within individuals between corresponding 5 cycle increments
(e.g.~15 to 20, 20 to 25, etc.). Conversely, this may be due to low
power and on observation there does seem to be a trend that 20x and 25x
communities are very different {[}Figure 3{]}. This finding, in
conjunction with the PERMANOVA results, suggest that cycle number can
change bacterial community calculations but that these differences are
minimal once at least 25 cylces are reached. Increasing the sub-sampling
depth, for some DNA polymerase may reduce some of these observed
community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence
differences between the different HiFi DNA polymerases {[}Figure 5{]}.
The chimera prevalence was very strongly correlated with the number of
OTUs and this value is relied upon heavily for many different downstream
community metric calculations. However, Bray-Curtis analysis with
PERMANOVA showed few differences based on DNA polymerase. Since it is
possible that many of the increased number of OTUs, generated as cycle
number increases, are not highly abundant that the Bray-Curtis index is
able to successfully downweight these respective OTUs (Minchin, 1987).
So, choice of downstream diversity metric could be an important
consideration to mitigate these observed changes due to high chimera
prevalence in HiFi DNA polymerase such as Platinum.

Our observations suggest that there are clear HiFi dependent differences
in both per base error rate and chimeras that cannot be removed using
bioinformatic approaches {[}Figure 4 \& 5{]}. Although it may be a
natural assumption that the variation may be due to the DNA polymerase
family, the highest chimera rate, from Platinum, was a family A
polymerase while the lowest, from Accuprime, was also an A polymerase
(Ishino \& Ishino, 2014). In fact, from the material safety data sheet
(MSDS), it is not clear what the difference between the two different
mixes really is. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate since all samples were also sequenced at
the same time as well as amplified using the same machine.

@@

\hypertarget{ref-polymerase_Ishino_2014}{}
Ishino S., Ishino Y. 2014. DNA polymerases as useful reagents for
biotechnology â the history of developmental research in the field.
\emph{Frontiers in Microbiology} 5. DOI:
\href{https://doi.org/10.3389/fmicb.2014.00465}{10.3389/fmicb.2014.00465}.

@@

\textbf{Figure 1: Normalized Fecal Number of OTUs.} A) Sub-sampled to
1000 reads. B) Sub-sampled to 5000 reads. C) Sub-sampled to 10000 reads.
D) Sub-sampled to 15000 reads. The dotted line represents no change from
the mean number of OTUs within that specific individual.

\textbf{Figure 2: Mock Sample Variability in Number of OTUs based on
HiFi DNA Polymerase.} A) Sub-sampled to 1000 reads. B) Sub-sampled to
5000 reads. C) Sub-sampled to 10000 reads. The dotted line represents
the number of OTUs genereated when the mock reference sequences are run
through the pipeline.

\textbf{Figure 3: Five Cycle Interval Community Differences.} A) Fecal
samples sub-sampled to 1000 reads. B) Fecal samples sub-sampled to 5000
reads. C) Fecal samples sub-sampled to 10000 reads. D) Mock samples
sub-sampled to 1000 reads. E) Mock samples sub-sampled to 5000 reads. F)
Mock samples sub-sampled to 10000 reads. The solid black lines represent
the median Bray-Curtis index difference within sample for each 5 cycle
interval.

\textbf{Figure 4: HiFi DNA Polymerase Per Base Error Rate in Mock
Samples.} A) Error rate before the merger of sequences with pre.cluster
and the removal of chimeras with VSEARCH. B) Error rate before the
removal of chimeras with VSEARCH. C) Full pipeline. The error bars
represent the 75\% interquartile range of the median.

\textbf{Figure 5: HiFi DNA Polymerase Chimera Prevalence in Mock
Samples.} A) Chimera sequence percentage before the merger of sequences
with pre.cluster and the removal of chimeras with VSEARCH. B) Chimera
sequence percentage before the removal of chimeras with VSEARCH. C) Full
pipeline. The error bars represent the 75\% interquartile range of the
median.

\textbf{Figure 6: The Correlation between Number of OTUs and Chimeras.}
A) Correlation before the merger of sequences with pre.cluster and the
removal of chimeras with VSEARCH. B) Correlation before the removal of
chimeras with VSEARCH. C) Correlation with full pipeline.

@@

\textbf{Figure S1: HiFi DNA Polymerase Sequence Error Prevalence in Mock
Samples.} A) Sequence error prevalence before the merger of sequences
with pre.cluster and the removal of chimeras with VSEARCH. B) Sequence
error prevalence before the removal of chimeras with VSEARCH. C) Full
pipeline. The error bars represent the 75\% interquartile range of the
median.

\textbf{Figure S2: HiFi DNA Polymerase Nucleotide Subsitutions in Mock
Samples.}
